---
title: "salmon_survival"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{salmon_survival}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup}
library(pssalmonsurvival)

```

Map of model extent

```{r map model ectent, echo=TRUE}

file.name <- "amps_model_map.png"

make_map(file.name)
```


Model ensemble was developed with alternate parametrizations. Plot biomass trajectories based on simulations that run the model ensemble 30 years under constant fishing rates. We only use 6 model variants, 2:5 and 7:8

```{r}

data("ensemblebiomass")
plotmodels <- c(1,6) # eliminated model versions 1 & 6

plot_ensemblebiomass(ensemblebiomass, plotmodels)

```


Plot salmon survival for model ensemble, survival is defined as the proportion of age 1 salmon that survive to age 5, the cohort is lagged over time. 

```{r}
data("ensemblenumbersage")
data("salmongroups")
plotmodels <- c(1,6) # eliminated model versions 1 & 6

plot_ensemble_survival(ensemblenumbersage, salmongroups, plotmodels)

```

