---
title: "salmon_survival"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{salmon_survival}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Package for the article
Effect of multiple pressures on early marine survival of juvenile salmon in Puget Sound

Hem Nalini Morzaria-Luna, I.C. Kaplan, C.J. Harvey, M. Schmidt, E.A. Fulton, R. Girardin, and P. MacCready


```{r setup}
install.packages("rnaturalearthhires", repos = "http://packages.ropensci.org", type = "source")
library(pssalmonsurvival)

```

Map of model extent in Puget Sound

```{r map model extent, echo=TRUE}

file.name <- "amps_model_map.png"

make_map(file.name)

```

Plot food web

```{r}
#Plot model food web

data("ppreymatrix")
plot.name <- "ps_foodweb.png"
  
plot_foodweb(ppreymatrix, plot.name)


```

Model ensemble was developed with alternate parametrizations. Plot biomass trajectories based on simulations that run the model ensemble 30 years under constant fishing rates. We only use 6 model variants, 2:5 and 7:8
Model biomass comes from 30-year runs of the base models. There are 8 model variants that differ in invertebrate growth rate and vertebrate density dependence 

```{r plot biomass base runs, echo=FALSE}

#biomass comparison between model variants, base runs
data("ensemblebiomass")
plotmodels <- c(1,6) # eliminated model versions 1 & 6

plot_ensemblebiomass(ensemblebiomass)

```


Plot salmon survival for model ensemble, survival is defined as the proportion of age 1 salmon that survive to age 5, the cohort is lagged over time. 

```{r plot survival base runs, echo=FALSE}
#numbers at age from base runs
data("ensemblenumbersage")
data("salmongroups")
plotmodels <- c(1,6) # eliminated model versions 1 & 6

plot_ensemble_survival(ensemblenumbersage, salmongroups)

```

Make forcings for model simulations, by scenario

```{r make scenario forcings, include=FALSE}

#groups for each scenario
sim.groups <- list(c("Pinniped", "California_sea_lion", "Harbor_seal"), 
                c("PinkSY_Fish", "ChumFSY_Fish"), 
                c("Herring_PS", "Herring_Cherry"),
                "Porpoise",
                "Spinydog_Fish", 
                c("Pisc_Seabird","NonPisc_Seabird"),
                "Gel_Zoo", 
                c("ChinookY_Fish","ChinookSY_Fish","CohoHY_Fish","ChumHSY_Fish"),
                #chinook hatchery are repeated twice for a scenario with just Chinook
                c("ChinookY_Fish_2","ChinookSY_Fish_2"))

data("functionalgroups")
data("salmongroups")

ncfile = "AMPS.nc"
num.forcing.years = 50
bash.file = "amps_cal.sh"
force.prm = "PugetSound_force.prm"
func.groups = functionalgroups

#scenario 20% decrease

rate.multiplier = 0.8

lapply(sim.groups, make_forcing, ncfile, func.groups, num.forcing.years, rate.multiplier, bash.file, force.prm, salmongroups)

#scenario 20% increase

rate.multiplier = 1

lapply(sim.groups, make_forcing, ncfile, func.groups, num.forcing.years, rate.multiplier, bash.file, force.prm, salmongroups)

rate.multiplier = 1.2

lapply(sim.groups, make_forcing, ncfile, func.groups, num.forcing.years, rate.multiplier, bash.file, force.prm, salmongroups)


```


Plot salmon survival for scenario results, survival is defined as the proportion of age 1 salmon that survive to age 5, the cohort is lagged over time. 

```{r plot survival scenarios, echo=FALSE}

#base case survival for each model variant created above

base.survival <- readr::read_csv("base_survival.csv") %>% 
  tidyr::drop_na() %>%
  dplyr::mutate(max_year = max(year_no)) %>% 
  dplyr::filter(year_no<=(max_year-3)) %>% 
  dplyr::filter(year_no==max(year_no))

#numbers at age from base runs
data("ensemblenumbersagescenarios")
data("salmongroups")
#plotmodels <- c(1,6) # eliminated model versions 1 & 6

plot_ensemble_survival_scenarios(ensemblenumbersagescenarios, salmongroups, plotmodels, base.survival)

```

